---
title: "Intro to cstime"
output: rmarkdown::html_vignette
author: "Chi Zhang"
date: "2021-09-15"
vignette: >
  %\VignetteIndexEntry{Intro to cstime}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`cstime` provides date and time functions for public health surveillance. Its main purpose is consistent conversion between:

- Date.
- ISO year, ISO week, ISO yearweek.
- Season week.

`cstime` is part of the [csverse](https://niphr.github.io/packages.html) package suite.

```{r setup}
library(cstime)
```

## What cstime is for

Public health surveillance counts cases by week. The week is the unit of analysis, not the calendar month. `cstime` converts a date into the ISO week, ISO year, season or quarter that contains it, and converts back the other way.

## How the function names work

Almost every name is built the same way:

```
<from>_to_<to>_<suffix>
```

The suffix says what type comes back. `_c` returns a character string, `_n` returns a number, `_date` returns a `Date`.

```{r}
date_to_isoyearweek_c("2021-01-01")
isoyearweek_to_isoyear_n("2021-01")
isoyearweek_to_last_date("2021-01")
```

cstime exports 33 functions. All 33 end in `_c` (17 of them), `_n` (13) or `_date` (3). All but one also follow `<from>_to_<to>`. The exception is `now_c()`, which reads the clock instead of converting. Once you know the pattern, you can guess a function name instead of searching for it.

## Three things that are easy to get wrong

**An ISO year is not a calendar year.** 2020-12-31 and 2021-01-01 sit in different calendar years, but both sit in ISO yearweek 2020-53.

```{r}
d <- as.Date(c("2020-12-31", "2021-01-01"))
data.frame(
  date = d,
  calyear = date_to_calyear_n(d),
  isoyear = date_to_isoyear_n(d),
  isoyearweek = date_to_isoyearweek_c(d)
)
```

**An ISO year has 52 or 53 weeks.** Do not assume 52. 2015, 2020 and 2026 each have 53.

```{r}
data.frame(
  isoyear = 2015:2026,
  weeks_in_isoyear = isoyear_to_last_isoweek_n(2015:2026)
)
```

**A season week is not an ISO week.** A season starts in ISO week 35 and runs past New Year, so one season covers two ISO years. A season also numbers its own weeks from 1, so ISO week 35 is season week 1.

```{r}
d <- as.Date(c("2020-08-24", "2020-12-28", "2021-01-04", "2021-08-29"))
data.frame(
  date = d,
  isoyearweek = date_to_isoyearweek_c(d),
  season = date_to_season_c(d)
)

isoweek_to_seasonweek_n(c(35, 52, 1, 34))
```

## cstime and the other cs packages

`cstime` needs no other cs package to run. `cstidy` needs cstime: when cstidy fills the time columns of a surveillance table -- isoyear, isoweek, season, quarter -- those values come from cstime.

## Where to go next

- [Date, year, week conversion](date_conversion.html) covers dates, ISO years and ISO weeks.
- [Season week](season.html) covers season week numbering, including what happens in ISO week 53.

## Date to ISO year

```{r}
date_to_isoyear_c('2021-01-01')
```

## ISO yearweek string to ISO year and ISO week

```{r}
isoyearweek_to_isoyear_c("2021-02")
isoyearweek_to_isoweek_c("2021-02")
```

## Season week and ISO week

```{r}
seasonweek_to_isoweek_n(10)
isoweek_to_seasonweek_n(1)
```
