---
title: "Intro to cstime"
output: rmarkdown::html_vignette
author: "Chi Zhang"
date: "2021-09-15"
vignette: >
  %\VignetteIndexEntry{Intro to cstime}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`cstime` provides date and time functions for public health surveillance. Its main purpose is consistent conversion between:

- Date.
- Isoyear, isoweek, isoyearweek.
- Season week.

It also provides functions to generate rolling date ranges for computing weekly, bi-weekly, and 4-week averages.

`cstime` is part of the [csverse](https://niphr.github.io/packages.html) package suite.

```{r setup}
library(cstime)
library(magrittr)
```

## Date to isoyear

```{r}
date_to_isoyear_c('2021-01-01')
```

## Isoyearweek string to isoyear and isoweek

```{r}
isoyearweek_to_isoyear_c("2021-02")
isoyearweek_to_isoweek_c("2021-02")
```

## Season week and isoweek

```{r}
seasonweek_to_isoweek_n(10)
isoweek_to_seasonweek_n(1)
```
