---
title: "Season week"
author: "Chi Zhang"
date: "2022-03-14"
output:
  rmarkdown::html_vignette:
    fig_width: 6
    fig_height: 6
vignette: >
  %\VignetteIndexEntry{Season week}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Season week is the week number within an epidemiological season. It is used for surveillance outcomes such as influenza, where aligning data to a season rather than a calendar year makes trends easier to compare across years.

Isoweek runs from 1 to 53; season week runs from 1 to 52. When isoweek is 53, the corresponding season week is 18.5.

```{r}
library(cstime)
library(magrittr)
library(data.table)
```

## Mapping between isoweek and season week

The chart below shows the relationship across all isoweek values, with the special case of isoweek 53 highlighted.

```{r, echo = FALSE}
library(ggplot2)
pd <- data.table(isoweek = 1:53, seasonweek = isoweek_to_seasonweek_n(1:53))
q <- ggplot(pd, aes(x=seasonweek, y = isoweek))
q <- q + geom_point()
q <- q + geom_point(data = pd[isoweek==53], mapping = aes(color="Isoweek==53"))
q <- q + scale_color_discrete(NULL)
q <- q + scale_x_continuous("seasonweek", breaks = seq(1,52,4))
q <- q + scale_y_continuous("isoweek", breaks = seq(1,53,4))
q
```

## Conversion functions

```{r}
seasonweek_to_isoweek_c(10)
seasonweek_to_isoweek_n(10)
isoweek_to_seasonweek_n(1)
seasonweek_to_isoweek_n(1:52)
isoweek_to_seasonweek_n(1:53)
```


